---
title: "Temperature dependent development"
author: "Tyson Wepprich"
date: "November 16, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(devRate)
library(stringr)
library(lubridate)

```

## Goals
Analyze three lab studies of Aphalara development and survival rates at different temperature treatments. Compare linear and nonlinear temperature dependence of growth rates. Possibly combine the three studies' results for optimal physiological parameter estimates.

## The data
First, I need to transform data for analysis from different formats in 3 studies.

```{r data munging}
# myint
rawdat <- read.csv("data/dev_myint.csv", header = TRUE)

# combine male and females
# just do full development from egg to emerging adult
# simulate individual data?
myint_dev <- rawdat %>% 
  group_by(sex, temperature) %>% 
  do(simdata = rnorm(n = .$n, mean = .$total_mean, sd = .$total_se * sqrt(.$n))) %>% 
  unnest() %>% 
  mutate(time1 = round(simdata),
         temp2 = temperature,
         time2 = 0,
         n = 1,
         censored = 0)

# shaw
fs <- list.files(path = "data", full.names = TRUE)
fs <- fs[grep(pattern = "shaw", x = fs, fixed = TRUE)]
shaw_list <- list()
for (i in fs){
  tmp <- read_csv(i, col_names = FALSE)
  startdate <- str_split_fixed(tmp[1, 2], pattern = "-", 2)[2]
  startdate <- dmy(sub(pattern = ".", replacement = "/", x = startdate, fixed = TRUE))
  temperature <- as.numeric(str_split_fixed(string = str_split_fixed(string = i, pattern = "_", n = 3)[3], pattern = ".csv", n = 2)[1])
  
  nms <- tmp[3, ] %>% mutate_all(as.character)
  dat <- tmp[-c(1:3), ]
  names(dat) <- nms
  dat <- dat %>% 
    gather(key = enddate, value = n, 3:ncol(.)) %>% 
    mutate(enddate = dmy(enddate),
           time1 = enddate - startdate,
           temperature = temperature)
  shaw_list[[length(shaw_list) + 1]] <- dat
}
shaw_dev <- bind_rows(shaw_list) %>%
  rename(neggs = `number of eggs`) %>% 
  mutate(n = as.numeric(n),
         time1 = as.numeric(time1)) %>% 
  filter(complete.cases(.))
shaw_dev <- shaw_dev[rep(seq(nrow(shaw_dev)), shaw_dev$n), c("rep", "time1", "temperature")]


# temperature transfer
rawdat <- read_csv("data/dev_transfer.csv")
transfer_dev <- rawdat %>% 
  gather(key = enddate, value = n, `14/07/08`:ncol(rawdat)) %>% 
  mutate(time_all = as.numeric(dmy(enddate) - mdy(eggslaid)),
         temperature = tempto, 
         temp2 = 15,
         time1 = 30, 
         time2 = time_all - 30) %>% 
  filter(n > 0)
trans_dev <- transfer_dev[rep(seq(nrow(transfer_dev)), transfer_dev$n), 
                          c("plant", "run", "time1", "time2", "temperature", "temp2")]


# lots of variation in control 15C transfer emergence
plt <- ggplot(trans_dev, aes(x = time2)) + 
  geom_histogram() +
  # facet_wrap(~temperature, ncol = 1) +
  facet_grid(run~temperature) +
  theme_bw()
plt
```
## Analyze with devRate R package

```{r devRate}
# linear models
mod_lin_myint <- devRateModel(eq = lamb_92,
                              temp = myint_dev$temperature,
                              devRate = myint_dev$time1^-1,
                              startValues <- list(list(Rm = 0.05, Tmax = 30, To = 10), list(T1 = 4)))
devRatePlot(eq = lamb_92, nlsDR = mod_lin_myint, temp = myint_dev$temperature, devRate = myint_dev$time1^-1)

mod_lin_shaw <- devRateModel(eq = campbell_74,
                              temp = shaw_dev$temperature,
                              devRate = shaw_dev$time1^-1)
devRatePlot(eq = campbell_74, nlsDR = mod_lin_shaw, temp = shaw_dev$temperature, devRate = shaw_dev$time1^-1)



```

## Plot all 3 with linear models

```{r linear}

mean15 <- trans_dev %>% 
  filter(temperature == 15) %>% 
  mutate(time1 = time1 + time2,
         devRate = 1 / time1)
mean15 <- mean(mean15$devRate)

trans_dev <- trans_dev %>% 
  mutate(devRate = (1 - time2 * mean15) / time1,
         experiment = "transfer") %>%
  select(temperature, devRate, experiment)
shaw_dev <- shaw_dev %>% 
  mutate(devRate = 1 / time1,
         experiment = "shaw") %>% 
  select(temperature, devRate, experiment)
myint_dev <- myint_dev %>% 
  mutate(devRate = 1 / time1,
         experiment = "myint") %>% 
  select(temperature, devRate, experiment)

alldat <- bind_rows(trans_dev, shaw_dev, myint_dev)

rawplt <- ggplot(alldat, aes(x = temperature, y = devRate, color = experiment)) +
  geom_jitter(alpha = .3)
rawplt

```

## Linear predictions

```{r plot preds}
lindat <- alldat %>% 
  filter(temperature != 30)
mod <- lm(devRate ~ temperature * experiment, data = lindat)

newdat <- expand.grid(c("myint", "shaw", "transfer"), seq(0, 30, .1))
names(newdat) <- c("experiment", "temperature")

newdat$devRate <- predict.lm(mod, newdat)
newdat <- newdat %>% filter(devRate >= 0)

tempmod <- lm(devRate ~ temperature, data = lindat)
tempdat <- data.frame(temperature = seq(0, 30, .1))
tempdat$devRate <- predict.lm(tempmod, tempdat)
tempdat <- tempdat %>% filter(devRate >= 0)

test <- rawplt +
  geom_line(data = newdat, aes(x = temperature, y = devRate, group = experiment)) +
  geom_line(data = tempdat, aes(x = temperature, y = devRate), color = "black") +
  theme_bw() +
  ggtitle("Aphalara daily development rate vs constant temperatures")
test
ggsave(filename = "AphaDevRate.png", plot = test, device = "png", width = 6, height = 5)

```



## Analyze nonlinear
Trying to emulate Regniere et al 2012 SAS code.

```{r nonlinear}

library(nlme)


mod <- nlme(model = epsij ~ log(pnorm(q = epsij, 
                                      mean = (-0.5 * s_eps * s_eps), 
                                      sd = s_eps) - pnorm(q = epsm1,
                                                          mean = (-0.5 * s_eps * s_eps),
data <- myint_dev
regniere_negLL <- function(params){
  # sig_eps <- params[1]
  # sig_ups <- params[2]
  psi <- params[1]
  Tb <- params[2]
  Tm <- params[3]
  omega <- params[4]
  
  tau <- 1 / (psi * temperature * (temperature - Tb) * (Tm - temperature) ^ omega)
  
  lik <- 0
  for (i in 1:nrow(data)){
    for (j in 1:length(unique(temperature))){
      
      q <- pnorm(q = )
      
    }
  }
  
  
}


```

